---
title: "Project4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
  runtime: shiny
---

```{r setup, include=FALSE}
library(rsconnect)
library(tidyverse)
library(here)
library(readr)
library(flexdashboard)
library(shiny)
library(magrittr)
library(forcats)
library(waffle)
library(leaflet)
library(htmltools)
library(DT)
library(here)
library(lubridate)
```

```{r, eval=FALSE}
cars <- mtcars
```

```{r, eval=FALSE}
library(jsonlite)
library(httr)


if(!dir.exists(here("data"))) { dir.create(here("data")) }

# saves data only once (not each time you knit a R Markdown)
if(!file.exists(here("data","covid_data.RData"))) {
  
  covid_key <- Sys.getenv("COVID_ACT_NOW_API_KEY")

  ## build the URL
  base <- 'https://api.covidactnow.org/v2/states.timeseries'
  info_key <- '.json?apiKey='

  ## put it all together
  API_URL <- paste0(base, info_key, covid_key)
  
  raw_data <- GET(API_URL)
  covid_data <- fromJSON(rawToChar(raw_data$content), flatten = TRUE)
  save(covid_data, file = here("data","covid_data.RData"), compress='xz')
}
```

```{r}
load(here("data","covid_data.RData"))
```

```{r}
covid_data_actual <- covid_data %>% select(-metricsTimeseries, -cdcTransmissionLevelTimeseries,
                                           -communityLevelsTimeseries, -riskLevelsTimeseries)

dt <- covid_data_actual %>% select(-actualsTimeseries)

n = length(dt)

out = map(.x = covid_data_actual$actualsTimeseries,
      .f = function(.x, ...){
                       .x %>% 
          select(`positiveTests`, `negativeTests`, `newCases`, `date`) %>%  
          relocate(date, .before=positiveTests) %>% 
          mutate(positiveTests = replace_na(positiveTests, 0),
                 negativeTests = replace_na(negativeTests, 0),
                 newCases = replace_na(newCases, 0)) %>%
          mutate(pos_ratio = positiveTests/(positiveTests + negativeTests))
                     })
covid_data_actual <- covid_data_actual %>% 
  select(-county, -hsa, -hsaName, -lat, -long, -hsaPopulation, -lastUpdatedDate) %>% 
  mutate(actualsTimeseries=out)

```

```{r}

lm_summary <- covid_data_actual$actualsTimeseries %>%
  map(~ lm(newCases ~ pos_ratio, data = .x)) %>%
  map(summary) 

lm_summary_r_square <- map_dbl(.x = lm_summary, .f=function(.x, ...){
      unlist(.x)$r.squared
  })

lm_summary_coef <- covid_data_actual$actualsTimeseries %>%
  map(~ lm(newCases ~ pos_ratio, data = .x)$coefficients)

lm_summary_coefficient <- map_dbl(.x = lm_summary_coef, .f=function(.x, ...){
      unlist(.x)[2]
  })

lm_summary_intercept <- map_dbl(.x = lm_summary_coef, .f=function(.x, ...){
      unlist(.x)[1]
  })

covid_data_actual_out <- covid_data_actual %>% 
  select(country, state, population,cdcTransmissionLevel, starts_with("actuals.")) %>%
  select(-starts_with("actuals.hsa"), -starts_with("actuals.vaccinations"), -starts_with("actuals.vaccinesAdministered")) %>%
  mutate(fit_r_square = lm_summary_r_square,
                             fit_coef = lm_summary_coefficient,
                             fit_intercept = lm_summary_intercept)

plot_data <- covid_data_actual_out %>% 
  select(state, population, actuals.cases, actuals.deaths, actuals.vaccinesDistributed) %>%
  pivot_longer(cols = c(population, actuals.cases, actuals.deaths, actuals.vaccinesDistributed), names_to = "stats", values_to = "values")
```


About 
===================================

Column
-----------------------------------
###
**What is the purpose of this dashboard?**


This dashboard has two purposes:

1. Academic-wise, to allow me to practice with creation of interactive dashboards and presenting my data. 
2. Data-wise, to present the timeseries data on state-wise COVID infection numbers in relation to vaccine distribution and population size.  

**Data Source**

The data used in this dashboard is retrieved from [COVID Act Now API](https://apidocs.covidactnow.org/), 
using the [jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) and [httr](https://cran.r-project.org/web/packages/httr/index.html) package. 



The Data
===================================

Column
-----------------------------------
###

Presented here is a trimmed down version of the "states.timeseries" data retrieved from 
the COVID Act Now API. 

The original dataframe directly converted from JSON object 
returned by the API call has nested dataframes and lists within cells, making it difficult to visualize,
columns that contain nested dataframes and lists are removed in the trimmed version for presentation.

Each row reports the COVID stats for one state, such as total cases, deaths, vaccine distribution, positive test counts and more.
Most states come with a CDC level time series data (not presented in this interactive table) that contains daily case counts, positive test etc.  

```{r}


DT::renderDataTable({
  DT::datatable(covid_data_actual %>% select(-actualsTimeseries),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```

Static 1
===================================


Column
-----------------------------------
### Chart 1 
```{r}


plot_data %>% ggplot(aes(fill=stats, x=values, y=state)) + 
    geom_bar(position="dodge", stat="identity", orientation ="y") + 
  labs(title = "Counts on Cases, Deaths, Vaccination distribution and Total Population",
       subtitle = "Across U.S. states with Data",
       caption = "Generally Vaccination distribution farly exceeds the Population by 2 to 3 times") +
  ylab("State (2 Letter Abbreviation)") + xlab("Count")
```

Static 2 
===================================


Column {.tabset}
-----------------------------------
### Tab 1 
```{r}
covid_data_actual_out %>% 
  ggplot(aes(x=actuals.vaccinesDistributed/population, 
             y=actuals.cases/population,
             label=state,
             color=state)) +
  geom_point() + 
  geom_smooth(color="black", method = "lm") +
  geom_text(hjust=0, vjust=0, size=3) + theme(legend.position="none") +
  labs(title = "# Vaccine distrubtion/Population vs. Cases/Population",
       subtitle = "Negative Association",
       caption = "Seems like some weak negative correlation, so having vaccines help") +
  ylab("Cases/Population") + xlab("# Vaccine distrubtion/Population") 
``` 

### Tab 2
```{r}
covid_data_actual_out %>% 
  ggplot(aes(x=actuals.hospitalBeds.capacity/population, 
             y=actuals.deaths/population,
             label=state,
             color=state)) +
  geom_point() + 
  geom_smooth(color="black", method = "lm") +
  geom_text(hjust=0, vjust=0, size=3) + theme(legend.position="none") +
  labs(title = "Hospital Bed Capacity/Population vs. Deaths/Population",
       subtitle = "Negative Association",
       caption = "Seems like some weak negative correlation, interestingly more hospital beds correaltes with more deaths") +
  ylab("Deaths/Population") + xlab("Hospital Bed Capacity/Population") 
``` 

Interactive 1 
===================================

Column {.sidebar}
-----------------------------------
```{r}
selectInput("var", "State:", 
            covid_data_actual_out$state)
```

Column
----------------------------------

### Something here 

```{r}

renderPlot({
  st = input$var
  sub_df = (covid_data_actual %>% filter(state==st))[[9]][[1]]
  
  sub_df %>% filter(newCases != 0) %>% mutate(date = as_date(date)) %>% 
    ggplot(aes(x=date, y=newCases)) + geom_point() +
    ylab("Number of Infections") +
    xlab("Date") +
    labs(title = paste0("Timeseries data for infection in ", st))
  })
```

Interactive 2 
===================================

Column
----------------------------------
### Something here 

```{r}
sgeo = data.frame(state= state.abb, lat = state.center$y, lng = state.center$x)
sgeo = sgeo %>% inner_join(covid_data_actual_out, by="state")
```

```{r}
library(leaflet)
library(leaflet.minicharts)
tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"
basemap <- leaflet(width = "100%", height = "400px") %>% 
  addTiles(tilesURL)


```

```{r}

df <- sgeo %>% 
  mutate(idle_beds = actuals.hospitalBeds.capacity - actuals.hospitalBeds.currentUsageTotal,
         covid_patient = actuals.hospitalBeds.currentUsageCovid,
         other_patient = actuals.hospitalBeds.currentUsageTotal - actuals.hospitalBeds.currentUsageCovid)

colors <- c("#fcba50", "#cccccc", "#4fc13c")

basemap %>%
  addMinicharts(
    sgeo$lng, sgeo$lat,
    type = "pie",
    chartdata = df[, c("covid_patient", "other_patient", "idle_beds")], 
    colorPalette = colors, 
    width = 60 * sqrt(df$actuals.hospitalBeds.capacity) / sqrt(max(df$actuals.hospitalBeds.capacity)), transitionTime = 0
  )
```

Analysis 
===================================

Column
----------------------------------
### Wanted to test whether the recorded COVID-testing data reflects COVID incidence over a time series across states. 

```{r, echo=TRUE}
covid_data_append <- covid_data_actual %>% select(state, actualsTimeseries)

timeseries_list = covid_data_append$actualsTimeseries
state_list = covid_data_append$state

map2_out <- map2(.x = state_list,
                 .y = timeseries_list,
                     .f = function(.x, .y){
                       mutate(.y, state=.x)
                     })
covid_timeseries_long <- bind_rows(map2_out)

covid_timeseries_long <- covid_timeseries_long %>% 
  mutate(pos_ratio = replace_na(positiveTests, 0))

```

Column
----------------------------------

```{r, fig.width=10,fig.height=20}

  covid_timeseries_long %>% ggplot(aes(x=pos_ratio, y=newCases, color=`state`)) +
  geom_point() +
  geom_smooth(color="black", method = "lm") + facet_wrap(facets="state", nrow=11,ncol=5, scale="free") + 
  labs(title = "Positive Ratio and Same day incidence",
       subtitle = "Across States",
       caption = "The correaltion is very bad for all states, Incidence is often underestimated by positive ratio") +
  ylab("New Cases (Incidence)") + xlab("Proportion Tested Positive") 
```

It seems like if you test more people you are able to get better prediction of how bad the incidence is but with a very small marginal gain. 








